redo of https://vincentarelbundock.github.io/marginaleffects/

```{r}
library(MASS)
library(nnet)
library(tidyverse)
library(marginaleffects)
library(palmerpenguins)
```

```{r}
penguins
```


```{r}
penguins.1 <- lm(bill_depth_mm ~ bill_length_mm, data = penguins)
summary(penguins.1)
```

```{r}
predictions(penguins.1)
```

```{r}
predictions(penguins.1, variables = "bill_length_mm")
```

introduce species

```{r}
ggplot(penguins, aes(x = bill_length_mm, y = bill_depth_mm, colour = species)) + 
  geom_point() + geom_smooth(method = "lm")
```


```{r}
penguins.2 <- lm(bill_depth_mm ~ bill_length_mm + species, data = penguins)
summary(penguins.2)
```

```{r}
predictions(penguins.2, variables = c("bill_length_mm", "species"))
```

```{r}
predictions(penguins.2, variables = "species")
```

```{r}
predictions(penguins.2, variables = "bill_length_mm")
```

```{r}
comparisons(penguins.2)
```

```{r}
summary(comparisons(penguins.2))
```

```{r}
penguins %>% group_by(species) %>% 
  summarize(mean_depth = mean(bill_depth_mm, na.rm = TRUE),
            mean_length = mean(bill_length_mm, na.rm = TRUE))
```

```{r}
plot_cap(penguins.2, condition = c("bill_length_mm", "species"))
```



rats from d29

```{r}
my_url <- "http://www.utsc.utoronto.ca/~butler/d29/rat.txt"
rats <- read_delim(my_url, " ")
rats
```

```{r}
rats2 <- rats %>% mutate(status = factor(status))
status.1 <- glm(status ~ dose, family = "binomial", data = rats2)
```

```{r}
summary(status.1)
```

```{r}
predictions(status.1)
```

```{r}
plot_cap(status.1, condition = "dose")
```

```{r}
predictions(status.1, variables = "dose")
```


custom doses

```{r}
new <- datagrid(model = status.1, dose = seq(0.5, 4.5, 1))
new
predictions(status.1, newdata = new)
```

```{r}
comparisons(status.1)
```

bigger dataset

```{r}
my_url <- "http://www.utsc.utoronto.ca/~butler/d29/rat2.txt"
rat2 <- read_delim(my_url, " ")
##
## -- Column specification ----------------------------
## cols(
## dose = col_double(),
## lived = col_double(),
## died = col_double()
## )
rat2
```

```{r}
response <- with(rat2, cbind(lived, died))
response
rat2.1 <- glm(response ~ dose,
family = "binomial",
data = rat2
)
summary(rat2.1)
```

```{r}
comparisons(rat2.1)
```

```{r}
predictions(rat2.1)
```



```{r}
predictions(rat2.1, variables = "dose")
```



```{r}
plot_cap(rat2.1, condition = "dose")
```

```{r}
my_url <-
"http://www.utsc.utoronto.ca/~butler/d29/sepsis.txt"
sepsis <- read_delim(my_url, " ")
sepsis
```

```{r}
sepsis.1 <- glm(death ~ factor(shock) + factor(malnut) + factor(alcohol) + age +
factor(bowelinf),
family = "binomial",
data = sepsis
)
summary(sepsis.1)
```


```{r}
predictions(sepsis.1, variables = "malnut")
```

```{r}
plot_cap(sepsis.1, condition = c("age", "malnut"))
```

```{r}
predictions(sepsis.1, variables = c("shock", "age"))
```



```{r}
new <- datagrid(model = sepsis.1, malnut = c(0, 1), shock = 0:1, alcohol = 0:1)
predictions(sepsis.1, newdata = new)
```

```{r}
comparisons(sepsis.1) %>% summary()
```

```{r}
summary(marginaleffects(sepsis.1))
```

not *quite* the same

now try ordered or multinomial

```{r}
my_url <- "http://www.utsc.utoronto.ca/~butler/d29/miners-tab.txt"
freqs <- read_table(my_url)
freqs
```

```{r}
freqs %>%
pivot_longer(-Exposure, names_to = "Severity", values_to = "Freq") %>%
group_by(Exposure) %>%
mutate(proportion = Freq / sum(Freq)) -> miners
miners
```

```{r}
miners %>%
mutate(sev_ord = fct_inorder(Severity)) -> miners
miners
```

```{r}
sev.1 <- polr(sev_ord ~ Exposure,
weights = Freq,
data = miners
)
summary(sev.1)
```

```{r}
p <- predictions(sev.1, type = "probs", variables = "Exposure")
p
```

```{r}
plot_cap(sev.1, condition = "Exposure", type = "probs")
```

that doesn't work

```{r}
plot_cap(sev.1, condition = "Exposure", type = "probs", draw = FALSE) %>% 
  ggplot(aes(x = Exposure.x, y= predicted, colour = group)) +
  geom_line()
```

multinomial

```{r}
my_url <- "http://www.utsc.utoronto.ca/~butler/d29/mlogit.csv"
brandpref <- read_csv(my_url)
brandpref
```

```{r}
brandpref %>%
mutate(sex = factor(sex)) %>%
mutate(brand = factor(brand)) -> brandpref
```

```{r}
brands.1 <- multinom(brand ~ age + sex, data = brandpref)
```

```{r}
step(brands.1)
```

```{r}
predictions(brands.1, variables = c("age", "sex"), type = "probs")
```

```{r}
plot_cap(brands.1, condition = c("age", "sex"), type = "probs", draw = FALSE) %>% 
  rename(age = condition1, sex = condition2, brand = group) %>% 
  ggplot(aes(x = age, y = predicted, colour = brand, shape = sex)) + geom_point() + geom_line()
```

```{r}
brands.4 <- update(brands.1, . ~ . + age:sex)
```

```{r}
step(brands.4)
```

```{r}
plot_cap(brands.4, condition = c("age", "sex"), type = "probs", draw = FALSE)  %>% 
  rename(brand = group, age = condition1, sex = condition2) %>% 
  ggplot(aes(x = age, y = predicted, colour = brand, shape = sex)) +
  geom_point() + geom_line()
  
```

very little difference from model without interaction

ancova


```{r}
url <- "http://www.utsc.utoronto.ca/~butler/d29/ancova.txt"
prepost <- read_delim(url, " ")
prepost
```

```{r}
prepost.1 <- lm(after ~ before * drug, data = prepost)
summary(prepost.1)
```

or, better (really):

```{r}
drop1(prepost.1, test = "F")
```

predictions with interaction

```{r}
predictions(prepost.1, variables = c("drug", "before"))
```

gap between a and b is bigger at the end.

```{r}
plot_cap(prepost.1, condition = c("before", "drug"))
```

predictions of after score for drug A higher than those for drug B by more when before score higher.


interaction comes out

```{r}
prepost.2 <- update(prepost.1, . ~ . - before:drug)
drop1(prepost.2, test = "F")
```

both before score and drug are significant in predicting after score

now plot predictions

```{r}
plot_cap(prepost.2, condition = c("before", "drug"))
```

now the lines are exactly parallel (a consequence of no interaction in model).


sleep data

```{r}
my_url <- "http://www.utsc.utoronto.ca/~butler/d29/sleep.txt"
sleep <- read_delim(my_url, " ")
sleep
```

```{r}
sleep.1 <- lm(atst ~ age, data = sleep)
summary(sleep.1)
```

predictions

```{r}
predictions(sleep.1)
```

A prediction for each child.

```{r}
predictions(sleep.1, variables = "age")
```

predictions for a five-number summary of age

custom ages

```{r}
new <- datagrid(model = sleep.1, age = c(10, 5))
new
```

and then predict with these

```{r}
predictions(sleep.1, newdata = new)
```

these are CIs for mean, as are these.

```{r}
plot_cap(sleep.1, condition = "age")
```

but we have to do prediction intervals the old way:

```{r}
p <- predict(sleep.1, newdata = new, interval = "p")
cbind(new, p)
```

visits to doctors


```{r}
my_url <-
"http://www.utsc.utoronto.ca/~butler/d29/regressx.txt"
visits <- read_delim(my_url, " ")
visits
```

```{r}
visits.1 <- lm(timedrs ~ phyheal + menheal + stress, data = visits)
summary(visits.1)

```

residuals

```{r}
ggplot(visits.1, aes(x = .fitted, y = .resid)) + geom_point()
```

```{r}
ggplot(visits.1, aes(sample = .resid)) + stat_qq() + stat_qq_line()
```

vs x

```{r}
predictions(visits.1) %>% 
  mutate(residual = timedrs - predicted) %>% 
  pivot_longer(phyheal:stress) %>% 
  ggplot(aes(x = value, y = residual)) + geom_point() +
  facet_wrap(~name, ncol = 2, scales = "free")
```


skipping

```{r}
boxcox(timedrs+1~phyheal + menheal + stress, data = visits)
```


```{r}
visits.3 <- lm(log(timedrs + 1) ~ phyheal + menheal + stress,
data = visits)
summary(visits.3)
```

```{r}
visits.4 <- lm(log(timedrs + 1) ~ phyheal + stress, data = visits)
summary(visits.4)
```

predictions for stress (actually of log timedrs + 1)

```{r}
plot_cap(visits.4, condition = "stress")
```

or do again

```{r}
unt <- function(y) {
  exp(y) - 1
}
```


```{r}
plot_cap(visits.4, condition = "stress", draw = FALSE) %>% 
  rename(stress = condition1) %>%
  mutate(across(c(predicted, conf.low, conf.high), ~unt(.))) %>% 
  ggplot(aes(x = stress, y = predicted, ymin = conf.low, ymax = conf.high)) + 
  geom_line() + geom_ribbon(alpha = 0.3)
```


other variable `phyheal` held at mean value. Or plot both (without ribbons this time)

```{r}
plot_cap(visits.4, condition = c("stress", "phyheal"), draw = FALSE) %>% 
  rename(stress = condition1, phyheal = condition2) %>% 
  mutate(across(c(predicted, conf.low, conf.high), ~unt(.))) %>% 
  ggplot(aes(x = stress, y = predicted, colour = phyheal)) +
  geom_line()
```



```{r}
predictions(visits.4, variables = "stress")
```

